---
date: "`r format(Sys.time(), '%d %B, %Y')`"
title:  "PHSU-MCC Cumulative Grant Report"
subtitle: "`r params$start_date` to `r params$end_date`"
author: ""
output: 
  officedown::rdocx_document:
    # NB: R execution does not seem supported here, so include path 
    # relative to markdown document (typically in reports/ vs templates/).
    reference_docx: "../templates/u54_docx.docx"
    page_margins:
      bottom: 0.5
      top: 0.5
      right: 0.5
      left: 0.5
      header: 0.5
      footer: 0.5
      gutter: 0
    page_size:
      width: 8.5
      height: 11
      orient: "portrait"

knit: pgreportr::knit_docx_with_datestamp

# Markdown parameters. You can specify a REDCap export
# file (.r or .csv pair) or a
# URL for importing from a REDCap server live.
#
# The start/end date are self-explanatory for report.
params:
  redcap_data: "PartnershipTrackingG_R_2022-01-20_1005.r"
  start_date: "2006-09-26"
  end_date: "2022-02-20"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.cap = TRUE)
library(officedown)
library(officer)
library(dplyr)
library(pgreportr)


start_date<-params$start_date
end_date <- params$end_date
redcap_version_date<-stringr::str_extract(params$redcap_data,"(?<=_)\\d{4}-\\d\\d-\\d\\d(?=_)")
```

REDCap data version: `r redcap_version_date`

<!---BLOCK_TOC--->

\newpage


# Overview

The purpose of this report is to list and count U54-related grants submitted, funded, and pending within the reporting period of `r start_date` to `r end_date`. The list below indicates the specific criteria used to define each category that is counted and listed. Please refer to Table \@ref(tab:summarytable) for details on the definition of specific terms used within the criteria.


```{r load-redcap}
suppressMessages(
  grants <- pgreportr::import_redcap_data(here::here("data-raw",params$redcap_data))
)
```

```{r create-initial-tables}

grants <- grants %>%
  dplyr::arrange(`Submission Date`)

grants_submitted<-grants %>%
  pgreportr::filter_grants_submitted_between(start =params$start_date, end=params$end_date)

grants_funded<-grants %>%
  pgreportr::filter_grants_funded_between(start=params$start_date,end=params$end_date) %>%
  arrange(`Funding Start Date`)

grants_pending<-grants %>%
  pgreportr::filter_grants_pending_review() 
```


```{r build-summary-table}

# The goal of this tibble is:
# tag: Easy name to refer to row
# Description: Header text
# Count: Number of category
# Data: Tibble of grants meeting the category criteria
# Detail: Long text with in-depth description of the way in which the table is derived.
summary_table<-tibble::tribble(
    ~tag, ~Description, ~Data, ~Detail,
  "submissions",
  "Total Grant Submissions", 
  grants_submitted,
"
The total number of grants with Submission Date in the range of {params$start_date} to {params$end_date}
are listed.
",
  
  "esi-related submissions",
  "Total ESI-Related Grants Submitted", 
  grants_submitted %>% filter_grants_esi_related(),
"
The total number of grants submitted in the range of {params$start_date} to {params$end_date} that are
identified as ESI-Related. ESI-Related is defined as a grant for which an ESI member is listed (not
necessarily PI, etc).
",

 # NOTE: This is a repeated table since ESI-mentions are the same grants
 # as ESI-related. There is an override below for the counts.
 "num esi-related mentions",
 "Total Number of times an ESI was involved in a submission",
  grants_submitted %>% filter_grants_esi_related(),
"
The total number of times an ESI was involved in a grant submitted in the range of {params$start_date}
to {params$end_date}. Each participant of a grant submission is counted if the participant is an ESI.
The same ESI may be counted multiple times as a member of distinct grant submissions.
",
  
  "submissions not funded",
  "Total Grant Submissions Not Funded", 
  grants_submitted %>% pgreportr::filter_grants_not_funded(),
"
The total number of grants submitted in the range of {params$start_date} to {params$end_date}
that were subsequently not funded. Note this is not the number of grants that were determined
to be Not Funded within the date range, but rather the number of submissions during the period
that were at some point determined Not Funded.
",
  
  "grants funded",
  "Total Grants Funded", 
  grants_funded,
"
The number of grants that were funded (awarded) between {params$start_date} to {params$end_date}.
Note this is determined by the annotated project start date.
",
  
  "esi grants funded",
  "Total ESI-related Grants Funded", 
  grants_funded %>% pgreportr::filter_grants_esi_related(),
"
The number of grants funded between {params$start_date} to {params$end_date} which are ESI-related,
meaning that a ESI is named within the grant.
",
  
  "submissions pending",
  "Total Grant Submissions Pending Review",
  grants_pending,
"
The number of grant submissions that are currently pending review. That is, grants that are 
submitted but no final decision has been determined. Note that given the time window of the
report, there may be no grants pending review.
",
  
  "esi submissions pending",
  "Total ESI-Related Grant Submissions Pending Review", 
  grants_pending %>% pgreportr::filter_grants_esi_related(),
"
Total number of grant submissions that are currently pending review and that involve an ESI within 
the grant. 
",

  ) %>%
  dplyr::mutate(
    `Count` = ifelse(tag =="num esi-related mentions",
                     purrr::map_int(Data, pgreportr::count_esi_mentions_in_grant),
                     purrr::map_int(Data, nrow)
    ),
    `Count` = tidyr::replace_na(`Count`, "0")

  )

```


Table \@ref(tab:summarytable) shows the summary of grant categories provided in this report. For each category, a brief 
description is provided along with the count of number of matching grants. An extended description,
including some of the logic associated with the filtering, is provided as well.

```{r summary-table, tab.cap = "Summary of reported grant categories.", tab.id="summarytable"}
summary_table %>%
  dplyr::select(Description, Count, Detail) %>%
  dplyr::mutate(Detail =  
                  stringr::str_remove_all(
                    purrr::map_chr(Detail, glue::glue), "\n")) %>%
  flextable::flextable() %>%
  flextable::width(width=c(1.5,0.8,5)) %>%
  flextable::align(j=2, align="right") %>%
  flextable::bg(bg="#4F81BD", part="header") %>%
  flextable::color(color="#FFFFFF", part="header") 


```

\newpage

# Individual Grant Reports

All grant categories are provided below, with a detailed list of each grant that matches the category.

```{r generate-individual-reports, results='asis'}
# Output a flextable per row of the summary.
summary_table %>%
  dplyr::rowwise() %>%
  purrr::pwalk(function(tag, Description, Data, Detail, Count, header) {
    cat("## ", Description, ": ", Count, "\n\n")
    cat(glue::glue(Detail), "\n\n")
    flextable::flextable_to_rmd(
      pgreportr::style_grants_as_flextable_gamma(Data),
      bookdown=TRUE
    )
    cat("\\newpage")
  })

```

# Excel Summary Output

The raw data for this grant report is also exported to an Excel file :

```{r define-excel-filename, format='asis'}
excel_filename <- here::here(
  "delivery",
  glue::glue("Cumulative_Grant_Tracking_Tables_{lubridate::today()}.xlsx")
)

officer::ftext(basename(excel_filename), officer::fp_text(font.family="Courier New"))
```

This file can be used to verify the numbers or for further analysis. Note that the Excel 
output is in "raw" form, in which indicator variables are included for each condition (e.g., `is_grant_funded`).

```{r write-xls, eval=TRUE}
xls_table <- summary_table %>%
  dplyr::rowwise() %>%
  dplyr::mutate(text_output =  
                  list(pgreportr::style_grants_as_text_alpha(Data))) %>%
  dplyr::select(tag, text_output) %>%
  tibble::deframe() %>%
  writexl::write_xlsx(path = excel_filename)
```


