---
date: "`r format(Sys.time(), '%d %B, %Y')`"
title:  "PHSU-MCC Monthly Publication Report"
author: ""
output: 
  officedown::rdocx_document:
    # NB: R execution does not seem supported here, so include path 
    # relative to markdown document (typically in reports/ vs templates/).
    reference_docx: "../templates/u54_docx.docx"
    page_margins:
      bottom: 0.5
      top: 0.5
      right: 0.5
      left: 0.5
      header: 0.5
      footer: 0.5
      gutter: 0
    page_size:
      width: 8.5
      height: 11
      orient: "portrait"

knit: pgreportr::knit_docx_with_datestamp

# Markdown parameters. You can specify a REDCap export
# file (.r or .csv pair) or a
# URL for importing from a REDCap server live.
#
# The start/end date are self-explanatory for report.
params:
  uri: "https://redcap.psm.edu/api/"
  end_date: "2022-05-09"
  use_cache: FALSE
---
                            

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.cap = TRUE)
library(officedown)
library(officer)
library(dplyr)
library(pgreportr)
library(ggplot2)

redcap_version_date <- lubridate::today()

# Current Partnership Grant reporting year
pg_yr<-pgreportr::convert_date_to_pg_year(params$end_date)
start_date <- pgreportr::pg_start_date(pg_yr)


```

---
subtitle: "`r start_date` to `r params$end_date`"
---

REDCap data version: `r redcap_version_date`

<!---BLOCK_TOC--->

\newpage






```{r load-redcap}

if ( params$use_cache ) {
  pubs <- readRDS(here::here("data","cached_pubs.rds"))
} else {
  # NOTE:
  # If you have not stored the key yet, use the following command at the R prompt to store it.
  #
  # keyring::key_set("pgreportr-REDCap-Publications-API-token", username = NULL, prompt = "REDCap API Publications Token: ")
  #
  # Get API key
  token <- keyring::key_get("pgreportr-REDCap-Publications-API-token", NULL)
  suppressMessages(
    pubs <- pgreportr::import_publications(params$uri, token)
  )
  saveRDS(pubs, here::here("data","cached_pubs.rds"))
}
```


```{r create-initial-tables}

pubs <- pubs %>%
  pgreportr::filter_pubs_between(
    start = pgreportr::pg_start_date(pg_yr),
    end = params$end_date
  ) %>%
  dplyr::arrange(`Publication Date`)


```

```{r build-summary-table}


# The goal of this tibble is:
# tag: Easy name to refer to row
# Description: Header text
# Count: Number of category
# Data: Tibble of pubs meeting the category criteria
# Detail: Long text with in-depth description of the way in which the table is derived.
summary_table<-tibble::tribble(
    ~tag, ~Description, ~Data,~Detail,
  "publications",
  "Total Publications", 
  pubs,
"
The total number of publications with Publication Date in the range of {start_date} to {params$end_date} are listed.
",
  
  "esi-related publications",
  "Total ESI-Related Publications", 
  pubs %>% filter_pubs_esi_related(),
"
The total number of publications in the range of {start_date} to {params$end_date} that are
identified as ESI-Related. ESI-Related is defined as a publication for which an ESI member is listed (not
necessarily first/last author, etc).
",
  
  # NOTE: This is a repeated table since ESI-mentions are the same pubs
  # as ESI-related. There is an override below for the counts so ignore the filter below.
  "num esi-related mentions",
  "Total Number of times an ESI was involved in a publication",
  pubs %>% pgreportr::filter_pubs_esi_related(),
"
The total number of times an ESI was involved in a publication in the range of {start_date} to {params$end_date}. Each participant of a publication is counted if the participant is an ESI.
The same ESI may be counted multiple times as a member of distinct publications.
",
  
  
   ) %>%
    dplyr::mutate(
    `Count` = ifelse(tag =="num esi-related mentions",
                     purrr::map_int(Data, ~pgreportr::count_esi_mentions(.x, "authors")),
                     purrr::map_int(Data, nrow)
    ),
    `Count` = tidyr::replace_na(`Count`, 0)

  )
  
```

```{r dashboard-cards, fig.height=1.5, fig.width=7.25, eval=FALSE}

info_box <- function(title="Submitted", content="21", fill="blue", col="white") {

  ggplot(data.frame(x=c(0,1), y=c(0,1)), aes(x=x,y=y)) +
    theme_void() +
    ggtext::geom_textbox(
      label = glue::glue(
        "<br><span style = 'font-size:16pt'>{title}</span><br><br><span style = 'font-size:48pt'>{content}</span>"
      ), 
      x=0.5,y=0.5,
      width=1, height=1, 
      fill=fill,col=col,
      box.r=unit(0,"npc"),
      #size=30,
      valign=0.5, halign=0.5
    ) 
}

grant_status_color <- c("submissions"="#34789a","grants funded"="#5ea54b","submissions pending"="cornsilk","submissions not funded"="#9e627e")
grant_status_color_palette <- function(tag) grant_status_color[tag]

grant_card_list <- list(
  info_box("Submitted",
            dplyr::filter(summary_table, tag=="submissions") %>% dplyr::pull("Count"), 
           fill=grant_status_color["submissions"],col="white"),
  info_box("Funded", 
           dplyr::filter(summary_table, tag=="grants funded") %>% dplyr::pull("Count"), 
           fill=grant_status_color["grants funded"], col = "white"),
  info_box("Not Funded", 
           dplyr::filter(summary_table, tag=="submissions not funded") %>% dplyr::pull("Count"), 
           fill=grant_status_color["submissions not funded"], col="white"),
  info_box("Pending", 
           dplyr::filter(summary_table, tag=="submissions pending") %>% dplyr::pull("Count"), 
           fill=grant_status_color["submissions pending"], col="black")
)

  
```

```{r dashboard-barplot, fig.width=7.25, fig.height=6, eval=FALSE}
ptable <- summary_table %>%
  dplyr::slice(
    rev(match(c("submissions","grants funded", "submissions pending","submissions not funded"), summary_table$tag))
  ) %>%
  dplyr::mutate(Description =stringr::str_wrap(Description, 15)) %>%
  dplyr::mutate(Description = factor(Description, levels = Description))


grant_bar_chart <- ggpubr::ggbarplot(ptable, x = "Count",y="Description", ylab = "",
                  palette = grant_status_color, 
                  fill = "tag",
                  sort.val ="none", width=0.9) +
  theme(legend.position = "none",
        axis.text=element_text(size=8)) 


plot_piechart <- function(df, main="", show_legend=FALSE) {
  p <- ggplot(df, aes(x=1, y=value, fill=label)) + 
    geom_col() + 
    coord_polar(theta="y") +
    geom_text(
      aes(label = value),
      position = position_stack(vjust = 0.5),
      size=8
    ) + 
    theme_void() +
    scale_fill_brewer() +
    xlab("") + ylab("") +
    theme(legend.title = element_blank()) +
    ggtitle(main) +
    xlim(c(-0.5,1.5)) 

  if ( !show_legend) p <- p + theme(legend.position="none")
  
  p
   
}

make_pie_df <- function(.x, all, esi) {
  pl <- .x %>%
    dplyr::filter(tag %in% c(all, esi)) %>%
    dplyr::select(tag, Count) %>%
    tibble::deframe()
  
  tibble::tribble(
    ~label, ~value,
    "Non-ESI Related", pl[all]-pl[esi],
    "ESI Related", pl[esi] 
  )
}

grant_piechart_list <- list(
  plot_piechart(make_pie_df(summary_table, "submissions", "esi-related submissions"), main="Submissions"),
  plot_piechart(make_pie_df(summary_table, "submissions pending", "esi submissions pending"), main="Pending"),
  plot_piechart(make_pie_df(summary_table, "grants funded","esi grants funded"), main="Funded"),
  cowplot::get_legend(plot_piechart(make_pie_df(summary_table, "grants funded","esi grants funded"), main="Funded",show_legend=TRUE))

)

# Complicated dashboard
# Top is the cards (full width)
# Bottom left is the barplot
# Bottom right are the pie charts
cowplot::plot_grid(
  nrow = 2, ncol = 1, rel_heights = c(1,2),
  cowplot::plot_grid(nrow=1, plotlist = grant_card_list),
  cowplot::plot_grid(
    nrow = 1, ncol = 2, rel_widths = c(2,3),
    grant_bar_chart, 
    cowplot::plot_grid(
      nrow = 2, ncol = 2,
      plotlist = grant_piechart_list
    )
  )
)

```

\newpage
# Overview

The purpose of this report is to list and count U54-related publications within the U54 grant year `r pg_yr` (`r start_date` - 
`r params$end_date`). The list below indicates the specific criteria used to define each category that is counted and listed. Please refer to Table \@ref(tab:summarytable) for details on the definition of specific terms used within the criteria.

Table \@ref(tab:summarytable) shows the summary of categories provided in this report. For each category, a brief 
description is provided along with the count of number of matching publications. An extended description,
including some of the logic associated with the filtering, is provided as well.

\newpage
```{r summary-table, tab.cap = "Summary of reported publication categories.", tab.id="summarytable"}
summary_table %>%
  dplyr::select(Description, Count, Detail) %>%
  dplyr::mutate(Detail =  
                  stringr::str_replace_all(
                    purrr::map_chr(Detail, glue::glue), "\n"," ")) %>%
  flextable::flextable() %>%
  flextable::width(width=c(1.5,0.8,5)) %>%
  flextable::align(j=2, align="right") %>%
  flextable::bg(bg="#4F81BD", part="header") %>%
  flextable::color(color="#FFFFFF", part="header") 


```

\newpage








# Individual Publication Reports

All publication categories are provided below, with a detailed list of each publication that matches the category.

```{r generate-individual-reports, results='asis'}
# Output a flextable per row of the summary.
summary_table %>%
  dplyr::rowwise() %>%
  purrr::pwalk(function(tag, Description, Data, Detail, Count, header) {
    cat("## ", Description, ": ", Count, "\n\n")
    cat(glue::glue(Detail), "\n\n")
    flextable::flextable_to_rmd(
      pgreportr::style_pubs_as_flextable_gamma(Data),
      bookdown=TRUE
    )
    cat("\\newpage")
  })

```




# Excel Summary Output

The raw data for this grant report is also exported to an Excel file :

```{r define-excel-filename, format='asis'}
excel_filename <- here::here(
  "delivery",
  glue::glue("Monthly_Publications_Tracking_Tables_{lubridate::today()}.xlsx")
)

officer::ftext(basename(excel_filename), officer::fp_text(font.family="Courier New"))
```

This file can be used to verify the numbers or for further analysis. Note that the Excel 
output is in "raw" form, in which indicator variables are included for each condition (e.g., `is_grant_funded`).

```{r write-xls, eval=TRUE}
xls_table <- summary_table %>%
  dplyr::rowwise() %>%
  dplyr::mutate(text_output =  
                  list(pgreportr::style_pubs_as_text_alpha(Data))) %>%
  dplyr::select(tag, text_output) %>%
  tibble::deframe() %>%
  writexl::write_xlsx(path = excel_filename)
```


